---
title: "Regression"
author: "Anna Yeaton"
date: "Fall 2019"
output:
  html_document:
    df_print: paged
fontsize: 11pt
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = T)
```

# Lab Section

In this lab, we will go over regression. We will be using the caret package in R. https://topepo.github.io/caret/train-models-by-tag.html

# Perfomance Metrics 

## Residual 

Deviation of the observed value to the estimated value (sample mean)
$$residual=y_i - \hat{y_i}$$
where $\hat{y_i}$ is the estimated value

## Mean Squared Error (MSE)

$$MSE=\frac{1}{n}\sum_{i=1}^{n}(y_i - \hat{y_i})^2$$

## Root Mean Squared Error (RMSE)
Same units as original data.

$$RMSE=\sqrt{MSE}$$

## L2 regularization : Ridge regression. Regularize by adding the sum of the coefficients, squared, to the function. 

$$Ridge Regression=\sum_{i=1}^{n}(y_i - w_0 - \sum_{j=1}^{p}w_jx_{ij})^2 + \lambda\sum_{j=1}^p(w_j)^2$$

## L1 regularization : Lasso Regression. Regularize by adding the sum of the absolute value of the coefficients to the model. Coefficient estimates may be pushed to zero -- Lasso can perform variable selection

$$Lasso Regression=\sum_{i=1}^{n}(y_i - w_0 - \sum_{j=1}^{p}w_jx_{ij})^2 + \lambda\sum_{j=1}^p|w_j|$$


\newpage

### The broad steps of Machine learning in R. 

1. Split the data into training and test. Set test aside. 

2. Fit a good model to the training data. 

3. See how your model did on the training data.

4. Test how your model performs on the test data. 

# Regression

```{r, include=FALSE}
library(caret)
library(MASS)
library(ggplot2)
library(dplyr)
library(ggfortify)



#Mauna Loa CO2 concentrations
data(airquality)
```

```{r}
length(airquality$Temp)
```

1. Split data into training and test set (75% in train set, 25% in test set)
```{r}
train_regression <- createDataPartition(airquality$Wind, p=0.75, list = FALSE, times = 1)
air_train <- airquality[train_regression,]
air_test <- airquality[-train_regression,]

head(air_train)
```


### Linear Regression

* Assumes a linear relationship. 
* Independent variables should not be correlated (no mulitcollinearity)
* The number of observations should be greater than the number of independent variables.


$$RSS=\sum(y_i - \hat{y_i})^2$$
We will predict the response of the Temperature based on Wind. 

This is the data we will fit a linear model to. 
```{r}
ggplot(data = air_train) +
  geom_point(aes(x=Wind, y=Temp)) +
  theme_bw()
```

2. Create and fit a linear model to predict Temperature from Wind using the training set

```{r}
# develop model of temp onto wind, and make predictions based off training inputs
model_linr <- train(Temp~Wind, data = air_train, method = "lm")
model_linr

pred <- predict(model_linr, air_train)
```


3. Vizualize how your model performed on the train data by plotting the regression line on top of the train data points. 
```{r}
ggplot(data = air_train) +
  geom_point(aes(x=Wind, y=Temp)) +
  theme_bw() +
  ggtitle("Linear regression line plotted over training data") +
  geom_line(aes(x = Wind, y = pred), color = "blue")
```


4. Explore how the model performs on the test data. For Linear Regression:

* The residuals should be close to zero.
* There should be equal variance around the regression line (homoscedasticity).
* Residuals should be normally distributed.
* Independent variables and residuals should not be correlated.

4 a) See how the model performs on the test data
```{r}
pred_test <- predict(model_linr, air_test)

ggplot(data = air_test) +
  geom_point(aes(x=Wind, y=Temp)) +
  ggtitle("Linear regression line plotted over test data") +
  theme_bw() +
  geom_line(aes(x = Wind, y = pred_test), color = "red")
```

4 b) Look at the residuals. Are they close to zero?
```{r}
#look at the median residual value. Close to zero is best
#help(summary)
summary(model_linr)
```
**Median residual is 1.298, this is ok but would ideally be smaller**

4 c) Plot predicted temperature vs observed temperature. A strong model should show a strong correlation
```{r}
ggplot() +
  geom_point(aes(x=air_test$Temp, y=pred_test)) +
  ggtitle("Observed vs. predicted temperature using linear regression") +
  xlab("Observed test temperature") +
  ylab("Predicted test temperature") +
  theme_bw() +
  geom_abline(intercept = 0, slope = 1, color = "blue")
```
**The oberved and predicted values don't cluster tightly around the central line -- it's not a particularly strong model**



4 d) Visualize the predicted values in relation to the real data points. Look for homoscedasticity
```{r}
# Extract coefficients from the model
m_coef <- coef(model_linr$finalModel)
m_coef
pred_test <- predict(model_linr, air_test)


# plot the regression line on the predicted values
# plot the original test values
ggplot(data = air_test) +
  geom_point(aes(x=Wind, y=pred_test)) +
  geom_point(aes(x=Wind, y=Temp), color = "blue") +
  ggtitle("Predicted Temperatures and linear regression line") +
  ylab("Temp: Predicted (black) and Real (blue)") +
  theme_bw() +
  geom_line(aes(x = Wind, y = pred_test), color = "red")
```


4 e) Residuals should be normally distributed. Plot the density of the residuals
```{r}
#residuals_lin <- residuals(model_linr$finalModel)
plot(density(resid(model_linr)))
# ggplot(data=residvpredict) +
#   geom_density(aes(residual))
```
**Residuals are about evenly distributed with mean around 0**


4 f) Independent variables and residuals should not be correlated
```{r}
cor.test(air_train$Wind, resid(model_linr))
```
**P-value >> 0.05, so we cannot reject the null hypothesis that residuals and independent variables are independent**

### Linear Regression with Regularization

5. Create a linear model using L1 or L2 regularization to predict Temperature from Wind and Month variables. Plot your predicted values and the real Y values on the same plot. 
```{r}
# L2 regression:
ctrl =  trainControl(method = "boot", 15)
model_l1 <- train(Temp~Wind + Month, data = air_train, method = 'ridge', trControl = ctrl)
pred_test <- predict(model_l1, air_test)

ggplot() +
  geom_point(aes(x=air_test$Temp, y=pred_test)) +
  ggtitle("Observed vs. predicted temperature using ridge regression") +
  xlab("Observed test temperature") +
  ylab("Predicted test temperature") +
  theme_bw() +
  geom_abline(intercept = 0, slope = 1, color = "blue")

```



